G\"odel's God in Isabelle/HOL

### 2013-10-31 Formalizations/Isabelle/GoedelGodSession/document/root.tex
updates for FU talk

\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym,amsmath,amssymb,a4wide}

@@

\title{G\"odel's God in Isabelle/HOL}
\author{Christoph Benzm\"uller and Bruno Woltzenlogel Paleo}
\maketitle

@@

\bibitem{Nitpick}
J.C. Blanchette and T.~Nipkow.
\newblock Nitpick: A counterexample generator for higher-order logic based on a
  relational model finder.
\newblock In {\em Proc. of ITP 2010}, number 6172 in LNCS, pp. 131--146.
  Springer, 2010.

%\bibitem{Satallax}
%C.E. Brown.
%\newblock Satallax: An automated higher-order prover.
%\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117.
 % Springer, 2012.

@@

\bibitem{Isabelle}
T.~Nipkow, L.C. Paulson, and M.~Wenzel.
\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
\newblock Number 2283 in LNCS. Springer, 2002.

@@

\bibitem{sobel2004logic}
J.H. Sobel.
\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
\newblock Cambridge U. Press, 2004.

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym,amsmath,amssymb,a4wide}
\usepackage{graphicx,xcolor}

@@

\title{G\"odel's God in Isabelle/HOL}
\author{Christoph Benzm\"uller and Bruno Woltzenlogel Paleo}
\date{November 1, 2013}
\maketitle

%\noindent\colorbox{gray}{\includegraphics[width=.99\textwidth]{$HOME/GoedelGod/Talks/FU-Berlin/ScottsScriptGrab}} %$

\noindent\fcolorbox{gray}{white}{
\begin{minipage}{.96\textwidth}\small
\begin{itemize}
\item[A1] Either a property or its negation is positive, but not
  both:  \hfill 
  $\all \phi [P(\neg \phi) \biimp \neg P(\phi)]$ \\[-1.5em]
\item[A2] A property necessarily implied \\ by a
  positive property is positive: \phantom{b} \hfill 
  $\all \phi \all \psi [(P(\phi) \wedge \nec \all x [\phi(x)
  \imp \psi(x)]) \imp P(\psi)]$ \\[-1.5em]
\item[T1] Positive properties are possibly exemplified: \hfill $\all
  \varphi [P(\varphi) \imp \pos \ex x \varphi(x)]$ \\[-1.5em]
\item[D1] A \emph{God-like} being possesses all positive properties: \hfill
  $G(x) \biimp \forall \phi [P(\phi) \to \phi(x)]$ \\[-1.5em]
\item[A3]  The property of being God-like is positive: \hfill   $P(G)$ \\[-1.5em]
\item[C\phantom{1}] Possibly, God exists: \hfill $\pos \ex x G(x)$ \\[-1.5em]
\item[A4]  Positive properties are necessarily positive: \hfill 
  $\all \phi [P(\phi) \to \Box \; P(\phi)]$ \\[-1.5em]
\item[D2] An \emph{essence} of an individual is a property possessed by it \\ and necessarily implying any of its properties:
  \phantom{b} \hfill $\ess{\phi}{x} \biimp \phi(x) \wedge \all
  \psi (\psi(x) \imp \nec \all y (\phi(y) \imp \psi(y)))$ \\[-1.5em]
\item[T2]  Being God-like is an essence of any
  God-like being: \hfill $\all x [G(x) \imp \ess{G}{x}]$ \\[-1.5em]
\item[D3] \emph{Necessary existence} of an individual is \\ the necessary exemplification of all its essences: 
  \phantom{b} \hfill $\NE(x) \biimp \all \phi [\ess{\phi}{x} \imp \nec
  \ex y \phi(y)]$ \\[-1.5em]
\item[A5] Necessary existence is a positive property: \hfill $P(\NE)$ \\[-1.5em]
\item[T3] Necessarily, God exists: \hfill $\nec \ex x G(x)$ 
\end{itemize}
\end{minipage}
}
\vskip1em

@@

\bibitem{Nitpick}
J.C. Blanchette and T.~Nipkow.
\newblock Nitpick: A counterexample generator for higher-order logic based on a
  relational model finder.
\newblock In {\em Proc. of ITP 2010}, LNCS 6172, pp. 131--146.
  Springer, 2010.

\bibitem{Satallax}
C.E. Brown.
\newblock Satallax: An automated higher-order prover.
\newblock In {\em Proc. of IJCAR 2012}, LNAI 7364, pp. 111 -- 117.
  Springer, 2012.

@@

\bibitem{Isabelle}
T.~Nipkow, L.C. Paulson, and M.~Wenzel.
\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
\newblock LNCS 2283. Springer, 2002.

@@

\bibitem{sobel2004logic}
J.H. Sobel.
\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
\newblock Cambr. U. Press, 2004.

### 2013-10-31 Formalizations/Isabelle/GoedelGodSession/output/document/GoedelGod.tex

 The formalization presented here has been carried and formally verified in the Isabelle/HOL 
 proof assistant; for more information on this system see the textbook by Nipkow, 
 Paulson, and Wenzel \cite{Isabelle}. More recent tutorials on Isabelle can be found 
 at: \url{http://isabelle.in.tum.de}.

@@

 \item For consistency checking, the model finder \cite{Nitpick} has been employed.

@@

lifted to type $\sigma$. Further connectives could be introduced analogously. Definitions 
could be used instead of abbreviations.%

@@

Constant symbol \isa{P} (G\"odel's "Positive") is introduced.%

@@

\ \ \isanewline

@@

Next, the symbol \isa{G} for "God-like" is introduced and defined 

@@

all positive properties:).%

@@

\ \ \isanewline

@@

@@

\ A{\isadigit{3}}\ T{\isadigit{1}}\ \isacommand{by}\isamarkupfalse%
\ metis%

@@

Symbol \isa{ess} for "Essence" is introduced and defined as 

@@

and necessarily implying any of its properties.).%

@@

\ {\isacharparenleft}metis\ {\isacharparenleft}lifting{\isacharparenright}\ A{\isadigit{1}}b\ A{\isadigit{4}}\ G{\isacharunderscore}def\ ess{\isacharunderscore}def{\isacharparenright}%

@@

Symbol \isa{NE}, for "Necessary Existence", is introduced and

@@

existence of an individual is the necessary exemplification of all its essences.).%

@@

\ A{\isadigit{5}}\ C\ T{\isadigit{2}}\ sym\ G{\isacharunderscore}def\ NE{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ metis%

@@

\ T{\isadigit{4}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \isanewline

@@

\ T{\isadigit{1}}\ T{\isadigit{3}}\ sym\ G{\isacharunderscore}def\ \isacommand{by}\isamarkupfalse%
\ metis%

@@

>>>>

 The formalization presented here has been carried out and verified in the Isabelle/HOL 
 proof assistant; for more information on this system see the textbook by Nipkow, 
 Paulson, and Wenzel \cite{Isabelle}. More recent tutorials on Isabelle can be found 
 at: \url{http://isabelle.in.tum.de}.

@@

 \item For consistency checking, the model finder Nitpick \cite{Nitpick} has been employed.

@@

lifted to type $\sigma$. The lifted connectives are \isa{m{\isasymnot}}, \isa{m{\isasymand}}, \isa{m{\isasymRightarrow}},
\isa{{\isasymforall}}, \isa{{\isasymPi}}, and \isa{{\isasymexists}}. Further connectives could be introduced analogously. 
Definitions could be used instead of abbreviations.%

@@

Constant symbol \isa{P} (G\"odel's `Positive') is declared.%

@@

\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}\ \isanewline

@@

Next, the symbol \isa{G} for `God-like'  is introduced and defined 

@@

all positive properties).%

@@

\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}%

@@

\ %

@@

\ {\isacharparenleft}metis\ A{\isadigit{3}}\ T{\isadigit{1}}{\isacharparenright}%

@@

Symbol \isa{ess} for `Essence' is introduced and defined as 

@@

and necessarily implying any of its properties).%

@@

\ {\isacharparenleft}metis\ A{\isadigit{1}}b\ A{\isadigit{4}}\ G{\isacharunderscore}def\ ess{\isacharunderscore}def{\isacharparenright}%

@@

Symbol \isa{NE}, for `Necessary Existence', is introduced and

@@

existence of an individual is the necessary exemplification of all its essences).%

@@

\ {\isacharparenleft}metis\ A{\isadigit{5}}\ C\ T{\isadigit{2}}\ sym\ G{\isacharunderscore}def\ NE{\isacharunderscore}def{\isacharparenright}%

@@

\ C{\isadigit{2}}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymexists}\ G{\isacharbrackright}{\isachardoublequoteclose}\ \isanewline
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}leo{\isadigit{2}}{\isacharbrackright}%

@@

\ {\isacharparenleft}metis\ T{\isadigit{1}}\ T{\isadigit{3}}\ G{\isacharunderscore}def\ sym{\isacharparenright}%

@@

\ \ \isacommand{lemma}\isamarkupfalse%
\ True\ \isacommand{nitpick}\isamarkupfalse%
\ {\isacharbrackleft}satisfy{\isacharcomma}\ user{\isacharunderscore}axioms{\isacharcomma}\ expect\ {\isacharequal}\ genuine{\isacharbrackright}%
\ %
\isacommand{oops}\isamarkupfalse%
It has been critisized that G\"odel's ontological argument implies what is called the 
modal collapse. The prover Satallax \cite{Satallax} can indeed show this, but verification with 
Metis still fails.%
\ \ \isacommand{lemma}\isamarkupfalse%
\ MC{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}p\ m{\isasymRightarrow}\ {\isacharparenleft}{\isasymbox}\ p{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
\ T{\isadigit{2}}\ T{\isadigit{3}}\ ess{\isacharunderscore}def\ sym%
\ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}satallax{\isacharbrackright}%
\ %
\isacommand{oops}\isamarkupfalse%
